Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2019-05-16 GWLDCII.tex
minor wording change
made comments up through theore 3.11

Note that we could replace $q_r$ by another propagator $q'_{r}$ in $W_r$, as long as $p_{r+1}$ and $q_r$ are on opposite sides of $q_r'$: such a new $q'_{r}$ would still satisfy all of the induction hypotheses. Therefore without loss of generality we may assume that $q_r$ is the innermost propagator of length $\geq 4$ in $W_r$ which satisfies the induction hypotheses, i.e. we assume that there are no other propagators with $p_{r+1}$ on one side and $q_r$ on the other.

@@

\begin{figure}
\includegraphics[width=15cm]{4th_vertex}
\caption{The two possible configurations in which a propagator can contribute its $<_i$-maximal support vertex to $I_i$.}
\label{fig:no fourth vertex}
\end{figure}

@@

\begin{proof}
We will prove the result by induction on the number of propagators.  If $W$ has one propagator then the result is immediate.  Now suppose $W$ has more than one propagator.  Since non-supporting vertices have no effect on Algorithm~\ref{alg:put GN on WLD}, it suffices to prove the result for $W$ with $F(\emptyset) = \emptyset$.  Then by Lemma~\ref{lem sian}, $W$ has at least one of the four situations illustrated in Figure~\ref{fig 3 cases}.

@@

In $V$ there are no propagators inside $q$, so we see from Figure~\ref{fig V diagram} that
\[l, l+1, \ldots ,j \in J^{V}_q(j) \text{ (if $l$ exists) and }  j+1 \in J^{V}_q(j+1).\]
Note that $j+2 \not\in J^{V}_q(j+1)$ by Lemma~\ref{lem no fourth vertex}, so by the induction hypothesis we must have $J^{V}_q(j+1) = \{j+1\}$ and $j+2 \in J^{V}_q(i)$. Thus there exist vertices $d,e \in [j+2,i+1]$ with $d <e$, such that
\[J^{V}_q(i) = [j+2,d-1], \quad J^{V}_q(i+1) = [d,e-1], \quad J^{V}_q(j) = [e,j], \quad J^{V}_q(j+1) = \{j+1\},\]
and all intervals are non-empty.

@@

\textbf{Left two cases}: These cases behave essentially identically (except when $j$ or $j+1$ are not in the support of $p$, which can occur in the second case only; see below) so we handle the majority of the proof for these two cases simultaneously. Let $1\leq a\leq 3$ be the number of non-supporting vertices inside $q$ in $V$; so these vertices are $l, \ldots, l+a-1$.  Write $p=(m,m+2)$ where $m\in \{i, i+1, l\}$.  Note that $l, \ldots, l+a-1$ are all in the support of $p$.

We first calculate $I^{W}_w$ for a starting vertex ${w \in [n]\backslash \{l, l+1, \ldots,j,j+1\}}$. Note that $p$ has no effect on other propagators for starting vertices in this range, while the value of $I^{W}_w(p)$ depends on how soon $q$ is assigned to a vertex, i.e. on the value of $I^{V}_w(q)$. Thus, if $w\in J_q^{V}(i)$ then
    \[
    I_w^{W}(r) =  \begin{cases}
        \max\{i+1, m\} & \text{if } r=p \\
        I_{w}^{V}(r) & \text{if } r\neq p
      \end{cases} 
    \]
    while if $w\in J_q^{V}(i+1)$ or $w\in J_q^{V}(j)$ then
    \[
    I_w^{W}(r) =  \begin{cases}
        l & \text{if } r=p \\
        I_{w}^{V}(r) & \text{if } r\neq p
      \end{cases} 
      \]
We also need to understand $I^{W}_w$ for $w \in \{l,l+1,\ldots,j,j+1\}$. For the majority of these vertices, we use the following observation: if $p$ is the first propagator to be assigned a value by $I_w^{W}$, then the remainder of $I_w^{W}$ proceeds identically to the assignments of $I^{V}_{w+1}$.  Thus we have for any $0\leq b <a$
    \[
    I_{l+b}^{W}(r) = \begin{cases}
      l+b & \text{if } r=p\\
      I_{l+b+1}^{V}(r) & \text{if } r\neq p
    \end{cases}
    \]
Similarly, if $j$ is in the support of $p$, then we have
     \[
       I_j^{W}(r) = \begin{cases}
         j & \text{if $r=p$ and $j$ is in the support of $p$}\\
         I_{j+1}^{V}(r) & \text{if $r\neq p$ and $j$ is in the support of $p$}\\
       \end{cases}
       \]
If $j$ is not in the support of $p$, then we must be in the second case of Figure~\ref{fig 3 cases} with two vertices in the right hand region.  In this case, if we start the algorithm at $j$ we need to know whether there will be any unassigned propagators other than $p$ when we reach vertex $i$, so as to know what $p$ contributes.

Consider the WLD $X$ formed from $V$ by moving the second end of $q$ to the edge $j-1$ instead of $j$.  $X$ is still admissible since we have not decreased the support of any set of propagators, so the induction hypothesis applies to it as well.  Note that $J_{j}^{V}(r) = J_{j}^{X}(r)$ for all $r\neq p$ and $J_{j}^{X}(r) = J_{j+1}^{X}(r)$ for $r\neq q,p$.  Additionally $I_{j+1}^{X}(q) = i$ by the induction hypothesis applied to $X$, and so if we start at $j+1$ and assign propagators to vertices according to the algorithm, when we reach vertex $i$ in $X$ all propagators other than $q$ must have been assigned.  Therefore if we start at $j$ in $W$, we first assign $q$ to $j$ then proceed to assign as in $X$ starting at $j+1$, and hence when we get to $i$ the only remaining unassigned propagator is $p$. Therefore
       \[
       I_j^{W}(r) = \begin{cases}

       m & \text{if $r =p$ and $j$ is not in the support of $p$} \\
       I_{j}^{V}(r) & \text{if $r\neq p$ and $j$ is not in the support of $p$}\end{cases}
       \] 
Finally, we consider what happens when we start the algorithm at vertex $j+1$. If $j+1$ is in the support of $p$ then we can argue as above to get
\[
       I_{j+1}^{W}(r)  = \begin{cases}
         j+1 & \text{if $r=p$ and $j+1$ is in the support of $p$}\\
         I_{j+2}^{V}(r) & \text{if $r\neq p$ and $j+1$ is in the support of $p$}.
       \end{cases}
       \]
Now suppose $j+1$ is not in the support of $p$. If we start at $j+1$ we need to know whether there are any unassigned propagators supported on edge $i$ when we reach vertex $i$. We already know that $J_q^{V}(j+1) = \{j+1\}$; in particular this means that $q$ contributes $i$ in $I_{j+2}^{V}$.  However the construction of $I^{V}_{j+1}$ first associates $q$ to $j+1$ and then proceeds identically to $I^{V}_{j+2}$.  In particular if $i$ was assigned in $I^{V}_{j+1}$, then it would not be available to assign to $q$ in $I^{V}_{j+2}$ as all other propagators supported at $i$ in $V$ come before $q$.

Therefore $p$ is the only potentially unassigned propagator on edge $i$ when we reach vertex $i$ in $W$, and
    \[
    I_{j+1}^{W}(r)  = \begin{cases}
      m & \text{if  $r=p$ and $j+1$ is not in the support of $p$}\\
      I_{j+1}^{V}(r) & \text{if  $r\neq p$ and $j+1$ is not in the support of $p$}
    \end{cases}
    \]
We can now describe the intervals $J^{W}_r(*)$ for the first two cases of Figure~\ref{fig 3 cases}. For $r \neq p$ the intervals are clearly still cyclic and appear in the correct order, and we can assemble the intervals for the $J_p^{W}(*)$ as follows.
    \begin{itemize}
      \item 
    If $m=l$ then either $a=2$ (so $l+1=m+1$, $j=m+2$, and $j+1=m+3$) or $a=3$ (so $l+1=m+1$, $l+2=m+2$, $j=m+3$, and $j+1$ is not in the support of $p$), and in both cases
    \begin{gather*}
    J^{W}_p(m) = [m+4, m], \qquad  J^{W}_p(m+1) = \{m+1\}, \\ J^{W}_p(m+2) = \{m+2\}, \qquad  J^{W}_p(m+3) = \{m+3\},
    \end{gather*}
    which are nonempty and otherwise as required.
  \item
        If $m=i+1$ then checking each of the three different possibilities for $a$ we likewise get
    \begin{gather*}
    J^{W}_p(m) = [m+4, d-1], \qquad  J^{W}_p(m+1) = [d, m+1], \\  J^{W}_p(m+2) = \{m+2\}, \qquad  J^{W}_p(m+3) = \{m+3\},
    \end{gather*}
    which are nonempty and otherwise as required.
  \item If $m=i$ then $a=1$ or $a=2$, in the former case $l=m+2$, $j=m+3$ and $j+1$ is not in the support of $p$ so
    \begin{gather*}
    J^{W}_p(m) = \{m+4\}, \qquad  J^{W}_p(m+1) = [m+5, d-1], \\  J^{W}_p(m+2) = [d, m+2], \qquad  J^{W}_p(m+3) = \{m+3\},
    \end{gather*}
    while in the latter $l=m+2$, $l+1=m+3$, and $j$ and $j+1$ are not in the support of $p$ so
    \begin{gather*}
    J^{W}_p(m) = [m+4, j+1], \qquad  J^{W}_p(m+1) = [j+2, d-1], \\  J^{W}_p(m+2) = [d, m+2], \qquad  J^{W}_p(m+3) = \{m+3\},
    \end{gather*}
    which are again as required.
    \end{itemize}

>>>>

Note that we could replace $q_r$ by another propagator $q'_{r}$ in $W_r$, as long as $p_{r+1}$ and $q_r$ are on opposite sides of $q_r'$: such a new $q'_{r}$ would still satisfy all of the induction hypotheses. Therefore without loss of generality we may assume that {\color{violet}$\cP_{in}(q_r)$ does not contain a propagator (other than $q_r$ itself)} \st{$q_r$ is the innermost propagator of length $\geq 4$ in $W_r$} which satisfies the induction hypotheses, i.e. we assume that there are no other propagators with $p_{r+1}$ on one side and $q_r$ on the other.

@@

\begin{figure}
\includegraphics[width=15cm]{4th_vertex}
\caption{The two possible configurations in which a propagator can contribute its $<_i$-maximal support vertex to $I_i$.}
\label{fig:no fourth vertex}
\end{figure}f

@@

\begin{proof}
We will prove the result by induction on the number of propagators.  If $W$ has one propagator then the result is immediate.  Now suppose $W$ has more than one propagator.  Since non-supporting vertices have no effect on Algorithm~\ref{alg:put GN on WLD}, it suffices to prove the result for $W$ {\color{violet} (possibly weakly admissible)} with $F(\emptyset) = \emptyset$.  Then by Lemma~\ref{lem sian}, $W$ has at least one of the four situations illustrated in Figure~\ref{fig 3 cases}.

@@

In $V$ there are no propagators inside $q$, so we see from Figure \ref{fig V diagram} that
\[l, l+1, \ldots ,j \in J^{V}_q(j) \text{ (if $l$ exists) and }  j+1 \in J^{V}_q(j+1).\]
Note that $j+2 \not\in J^{V}_q(j+1)$ by Lemma~\ref{lem no fourth vertex}, so  $J^{V}_q(j+1) = \{j+1\}$ and \hlfix{by the induction hypothesis we must have}{moved} $j+2 \in J^{V}_q(i)$. Thus there exist vertices $d,e \in [j+2,i+1]$ with $d <e$, such that
\ba J^{V}_q(i) = [j+2,d-1], \quad J^{V}_q(i+1) = [d,e-1], \quad J^{V}_q(j) = [e,j], \quad J^{V}_q(j+1) = \{j+1\}, \label{deeqs}\ea
and all intervals are non-empty.

@@

\textbf{Left two cases}: These cases behave essentially identically (except when $j$ or $j+1$ are not in the support of $p$, which can occur in the second case only; see below) so we handle the majority of the proof for these two cases simultaneously. \hlfix{Write $p=(m,m+2)$ where $m\in \{i, i+1, l\}$.}{moved} Let $1\leq a\leq 3$ be the number of non-supporting vertices inside $q$ in $V$ (note that there is at least 1); so these vertices are $l, \ldots, l+a-1$.  Note that these non-supporting vertices are all in the support of $p$.

We first calculate \hlfix{$I^{W}_w(?)$}{which prop(s), as a function?} for a starting vertex ${w \in [n]\backslash \{l, l+1, \ldots,j,j+1\}}$. Note that $p$ has no effect on other propagators for starting vertices in this range, {\color{violet}(if $r \neq p$, $I_w^W(r) = I_w^V(r)$)} while the value of $I^{W}_w(p)$ depends on \st{how soon $q$ is assigned to a vertex, i.e. on}\sanote{don't like the phrasing ``soon''} the value of $I^{V}_w(q)$. Thus, if $w\in J_q^{V}(i)$ then 
    \[
    I_w^{W}(r) =  \begin{cases}
        \max\{i+1, m\} & \text{if } r=p \\
        I_{w}^{V}(r) & \text{if } r\neq p
      \end{cases} 
    \]
    while if $w\in J_q^{V}(i+1)$ or $w\in J_q^{V}(j)$ then
    \[
    I_w^{W}(r) =  \begin{cases}
        l & \text{if } r=p \\
        I_{w}^{V}(r) & \text{if } r\neq p
      \end{cases} 
      \]
We also need to understand $I^{W}_w$ for $w \in \{l,l+1,\ldots,j,j+1\}$. For the \hlfix{majority}{$w \in [l, l+a-1]$} of these vertices, we use the following observation: if $p$ is the first propagator to be assigned a value by \hlfix{$I_w^{W}$, then the remainder of $I_w^{W}$}{algorithm instead of $I_w^W$?} proceeds identically to the assignments of $I^{V}_{w+1}$.  Thus we have for any $0\leq b <a$
    \[
    I_{l+b}^{W}(r) = \begin{cases}
      l+b & \text{if } r=p\\
      I_{l+b+1}^{V}(r) & \text{if } r\neq p
    \end{cases}
    \]
Similarly, if $j$ is in the support of $p$, then we have
     \[
       I_j^{W}(r) = \begin{cases}
         j & \text{if $r=p$ and $j$ is in the support of $p$}\\
         I_{j+1}^{V}(r) & \text{if $r\neq p$ and $j$ is in the support of $p$}\\
       \end{cases}
       \]
If $j$ is not in the support of $p$, then we must be in the second case of Figure~\ref{fig 3 cases} with two vertices in the right hand region.  In this case, if we start the algorithm at $j$ we need to know whether there will be any unassigned propagators other than $p$ when we reach vertex $i$, so as to know what $p$ contributes.

Consider the \st{WLD} {\color{violet} weakly admissible Wilson loop diagram}\sanote{actually, I think this is admissible, because we are removing a propagator from a weakly admissible. But it needs a few words?} $X$ formed from $V$ by \st{moving the second end of $q$ to the edge $j-1$ instead of $j$} {\color{violet} replacing the propagator $q = (i,j)$ with $q = (i, j-1)$}.  $X$ is still {\color{violet}weakly} admissible since we have not decreased the support of any set of propagators, so the induction hypothesis applies to it as well. Note that $J_{j}^{V}(r) = J_{j}^{X}(r)$ for all $r\neq p$ {\color{violet}and in particular for $r=q$. Further, } \st{and} $J_{j}^{X}(r) = J_{j+1}^{X}(r)$ for $r\neq q,p$.  Additionally $I_{j+1}^{X}(q) = i$ \st{by the induction hypothesis applied to $X$}, {\color{violet}because} \st{and so} if we start at $j+1$ and assign propagators to vertices according to the algorithm, when we reach vertex $i$ in $X$ all propagators other than $q$ must have been assigned.  Therefore if we start at $j$ in $W$, we first assign $q$ to $j$ then proceed to assign as in $X$ starting at $j+1$, and hence when we get to $i$ the only remaining unassigned propagator is $p$. Therefore
       \[
       I_j^{W}(r) = \begin{cases}

       m & \text{if $r =p$ and $j$ is not in the support of $p$} \\
       I_{j}^{V}(r) & \text{if $r\neq p$ and $j$ is not in the support of $p$}\end{cases}
       \] 
Finally, we consider what happens when we start the algorithm at vertex $j+1$. If $j+1$ is in the support of $p$ then we can argue as above to get
\[
       I_{j+1}^{W}(r)  = \begin{cases}
         j+1 & \text{if $r=p$ and $j+1$ is in the support of $p$}\\
         I_{j+2}^{V}(r) & \text{if $r\neq p$ and $j+1$ is in the support of $p$}.
       \end{cases}
       \]
Now suppose $j+1$ is not in the support of $p$. If we start at $j+1$ we need to know whether there are any unassigned propagators supported on edge $i$ when we reach vertex $i$. We already know that $J_q^{V}(j+1) = \{j+1\}$; \hlfix{in particular this means that $q$ contributes $i$ in $I_{j+2}^{V}$}{why? It is true but what does it follow from? The next few lines?}.  However the construction of $I^{V}_{j+1}$ first associates $q$ to $j+1$ and \hlfix{then proceeds identically to $I^{V}_{j+2}$.}{We use this fact a lot should it be a lemma?}  In particular if $i$ was assigned in $I^{V}_{j+1}$, then it would not be available to assign to $q$ in $I^{V}_{j+2}$ as all other propagators supported at $i$ in $V$ come before $q$.

Therefore $p$ is the only potentially unassigned propagator on edge $i$ when we reach vertex $i$ in $W$ {\color{violet} in the algorithm for $I_{j+1}^W$}, and
    \[
    I_{j+1}^{W}(r)  = \begin{cases}
      m & \text{if  $r=p$ and $j+1$ is not in the support of $p$}\\
      I_{j+1}^{V}(r) & \text{if  $r\neq p$ and $j+1$ is not in the support of $p$}
    \end{cases}
    \]
We can now describe the intervals $J^{W}_r(*)$ for the first two cases of Figure~\ref{fig 3 cases}. For $r \neq p$ the intervals are clearly still cyclic and appear in the correct order, and we can assemble the intervals for the $J_p^{W}(*)$ as follows.
    \begin{itemize}
      \item 
    If $m=l$ then either $a=2$ (so $l+1=m+1$, $j=m+2$, and $j+1=m+3$) or $a=3$ (so $l+1=m+1$, $l+2=m+2$, $j=m+3$, and $j+1$ is not in the support of $p$), and in both cases
    \begin{gather*}
    J^{W}_p(m) = [m+4, m], \qquad  J^{W}_p(m+1) = \{m+1\}, \\ J^{W}_p(m+2) = \{m+2\}, \qquad  J^{W}_p(m+3) = \{m+3\},
    \end{gather*}
    which are nonempty and otherwise as required.
  \item
        If $m=i+1$ then checking each of the three different possibilities for $a$ we likewise get
    \begin{gather*}
    J^{W}_p(m) = [m+4, d-1], \qquad  J^{W}_p(m+1) = [d, m+1], \\  J^{W}_p(m+2) = \{m+2\}, \qquad  J^{W}_p(m+3) = \{m+3\},
    \end{gather*}
    which are nonempty and otherwise as required. {\color{violet}Here, $d$ is as in display \eqref{deeqs}.}
  \item If $m=i$ then $a=1$ or $a=2$, in the former case $l=m+2$, $j=m+3$ and $j+1$ is not in the support of $p$ so
    \begin{gather*}
    J^{W}_p(m) = \{m+4\}, \qquad  J^{W}_p(m+1) = [m+5, d-1], \\  J^{W}_p(m+2) = [d, m+2], \qquad  J^{W}_p(m+3) = \{m+3\},
    \end{gather*}
    while in the latter $l=m+2$, $l+1=m+3$, and $j$ and $j+1$ are not in the support of $p$ so
    \begin{gather*}
    J^{W}_p(m) = [m+4, j+1], \qquad  J^{W}_p(m+1) = [j+2, d-1], \\  J^{W}_p(m+2) = [d, m+2], \qquad  J^{W}_p(m+3) = \{m+3\},
    \end{gather*}
    which are again as required.
    \end{itemize}
